Gray-Box Optimization using the Parameter-less Population Pyramid

### 2015-03-25 Gray Box Optimization using the Parameter-less Population Pyramid/main.tex
Some minor fixes + quartile bars on graphs

\numberofauthors{2} %
\begin{comment}
\author{
% 1st. author
\alignauthor
Brian W. Goldman\\
       \affaddr{BEACON Center for the Study of Evolution in Action}\\
       \affaddr{Michigan State University, U.S.A.}\\
       \email{brianwgoldman@acm.org}
% 2nd. author
\alignauthor
William F. Punch\\
       \affaddr{BEACON Center for the Study of Evolution in Action}\\
       \affaddr{Michigan State University, U.S.A.}\\
       \email{punch@msu.edu}
}
%\end{comment}
%\begin{comment}
\author{
% 1st. author
\alignauthor
Anonymous\\
       \affaddr{Group}\\
       \affaddr{Group}\\
       \affaddr{Organization}\\
       \email{email@site.com}
% 2nd. author
\alignauthor
Anonymous\\
       \affaddr{Group}\\
       \affaddr{Group}\\
       \affaddr{Organization}\\
       \email{email@site.com}
}
%\end{comment}

@@

\subsection{Novel r-order Subgraphs Algorithm}
\begin{figure}
  \begin{algorithmic}[1]
  \Procedure{ConnectedInducedSubgraphs}{}
    \State $closed \leftarrow \emptyset$
    \State $found \leftarrow []$
    \ForAll{$v \in V$}
      \State $closed \leftarrow closed \cup \{v\}$
      \State $found \leftarrow found + $\Call{CISG}{$v$, $\emptyset$, $closed$, $\emptyset$}
    \EndFor
    \State \Return $found$
  \EndProcedure
  \Procedure{CISG}{$v$, $subset$, $closed$, $open$}
    \State $subset' \leftarrow subset \cup \{v\}$
    \State $found \leftarrow [subset']$
    \If{$|subset'| \geq r$}
      \Return $found$
    \EndIf
    \State $closed\_here \leftarrow \emptyset$
    \State $open' \leftarrow open \cup adjacent(v)$
    \ForAll{$v' \in open'$ such that $v' \notin closed$}
        \State $closed\_here \leftarrow closed\_here \cup \{v'\}$
        \State $closed \leftarrow closed \cup \{v'\}$
        \State $recurse \leftarrow $\Call{CISG}{$v'$, $subset'$, $closed$, $open'$}
        \State $found \leftarrow found + recurse$
    \EndFor
    \State $closed \leftarrow closed - closed\_here$
    \State \Return $found$
  \EndProcedure
\end{algorithmic}
  \caption{Algorithm to recursively find all connected induced subgraphs of size $r$ or fewer.}
  \label{fig-connected-subgraphs}
\end{figure}

@@

\section{Tournament Uniform Crossover: TUX}
\label{sec-tux}
\begin{figure}
  \begin{algorithmic}[1]
  \Procedure{Iterate-TUX}{}
    \State Create random $solution$
    \State Hamming-Ball Hill Climb $solution$
    \ForAll{$T_i \in T$}
      \If{$T_i$ is empty}
        \State $T_i \leftarrow solution$
        \State \Return
      \EndIf
      \State Cross $solution$ with $T_i$ to create $2^{i+1}$ offspring
      \State Hamming-Ball Hill Climb each offspring
      \State $solution \leftarrow$ best of offspring, $solution$, and $T_i$
      \State $T_i \leftarrow$ empty
    \EndFor
    \State Add $solution$ to end of $T$
  \EndProcedure
\end{algorithmic}
  \caption{One iteration of TUX optimization. $T$ is an
           ordered list of solutions, each of which could be empty,
           awaiting a crossover partner.}
  \label{fig-TUX}
\end{figure}

@@

\section{Parameter-less Population Pyramid: P3}
\subsection{Origins}
\begin{figure}
  \begin{algorithmic}
  \Procedure{Iterate-P3}{}
    \State Create random solution
    \State Apply hill climber
    \If{solution $\notin hashset$}
      \State Add solution to $P_0$
      \State Add solution to $hashset$
    \EndIf

    \ForAll{$P_i \in pyramid$}
      \State Mix solution with $P_i$
      \If{solution's fitness has improved}
        \If{solution $\notin hashset$}
          \State Add solution to $P_{i+1}$
          \State Add solution to $hashset$
        \EndIf
      \EndIf
    \EndFor
  \EndProcedure
\end{algorithmic}
  \caption{One iteration of P3 optimization. $pyramid$ is an
           ordered set of populations and $hashset$ is a set
           of all solutions in $pyramid$.}
  \label{fig-p3}
\end{figure}

@@

For each problem class tested, we generated 50 instances.
Each method was run once on each instance, and limited to 3 hours of computation and 4 GB of memory.
Runs were performed on 2.5GHz Intel Xeon E5-2670v2 processors using the C++11 code available
from our website.\footnote{\url{https://github.com/name_of_repository/anonymous_name_of_project}}
Each time the run achieved a new best fitness we recorded the current amount of processing time used.
Timing includes the discovery of subgraphs to allow for comparison between different radius values.
When reporting the ``best'' fitness for an instance we mean the best fitness found by any method before
the time limit is reached. On all Nearest Neighbor NKq instances the ``best'' fitness
is also the global optimum, verified using dynamic programming. For Ising Spin Glass the ``best'' fitness
was the global optimum 44 of 50 times.

>>>>

\numberofauthors{2} %
%\begin{comment}
\author{
% 1st. author
\alignauthor
Brian W. Goldman\\
       \affaddr{BEACON Center for the Study of Evolution in Action}\\
       \affaddr{Michigan State University, U.S.A.}\\
       \email{brianwgoldman@acm.org}
% 2nd. author
\alignauthor
William F. Punch\\
       \affaddr{BEACON Center for the Study of Evolution in Action}\\
       \affaddr{Michigan State University, U.S.A.}\\
       \email{punch@msu.edu}
}
%\end{comment}
\begin{comment}
\author{
% 1st. author
\alignauthor
Anonymous\\
       \affaddr{Group}\\
       \affaddr{Group}\\
       \affaddr{Organization}\\
       \email{email@site.com}
% 2nd. author
\alignauthor
Anonymous\\
       \affaddr{Group}\\
       \affaddr{Group}\\
       \affaddr{Organization}\\
       \email{email@site.com}
}
%\end{comment}

@@

\subsection{Novel r-order Subgraphs Algorithm}
\begin{figure}
  \begin{algorithmic}[1]
  \Procedure{ConnectedInducedSubgraphs}{}
    \State $closed \leftarrow \emptyset$
    \State $found \leftarrow [~]$
    \ForAll{$v \in V$}
      \State $closed \leftarrow closed \cup \{v\}$
      \State $found \leftarrow found + $\Call{CISG}{$v$, $\emptyset$, $closed$, $\emptyset$}
    \EndFor
    \State \Return $found$
  \EndProcedure
  \Procedure{CISG}{$v$, $subset$, $closed$, $open$}
    \State $subset' \leftarrow subset \cup \{v\}$
    \State $found \leftarrow [subset']$
    \If{$|subset'| \geq r$}
      \Return $found$
    \EndIf
    \State $closed\_here \leftarrow \emptyset$
    \State $open' \leftarrow open \cup adjacent(v)$
    \ForAll{$v' \in open'$ such that $v' \notin closed$}
        \State $closed\_here \leftarrow closed\_here \cup \{v'\}$
        \State $closed \leftarrow closed \cup \{v'\}$
        \State $recurse \leftarrow $\Call{CISG}{$v'$, $subset'$, $closed$, $open'$}
        \State $found \leftarrow found + recurse$
    \EndFor
    \State $closed \leftarrow closed - closed\_here$
    \State \Return $found$
  \EndProcedure
\end{algorithmic}
  \caption{Algorithm to recursively find all connected induced subgraphs of size $r$ or fewer.}
  \label{fig-connected-subgraphs}
\end{figure}

@@

\section{Tournament Uniform\\ Crossover: TUX}
\label{sec-tux}
\begin{figure}
  \begin{algorithmic}[1]
  \Procedure{Iterate-TUX}{}
    \State Create random $solution$
    \State Hamming-Ball Hill Climb $solution$
    \ForAll{$T_i \in T$}
      \If{$T_i$ is empty}
        \State $T_i \leftarrow solution$
        \State \Return
      \EndIf
      \State Cross $solution$ with $T_i$ to create $2^{i+1}$ offspring
      \State Hamming-Ball Hill Climb each offspring
      \State $solution \leftarrow$ best of offspring, $solution$, and $T_i$
      \State $T_i \leftarrow$ empty
    \EndFor
    \State Add $solution$ to end of $T$
  \EndProcedure
\end{algorithmic}
  \caption{One iteration of TUX optimization. $T$ is an
           ordered list of solutions, each of which could be empty,
           awaiting a crossover partner.}
  \label{fig-TUX}
\end{figure}

@@

\section{Parameter-less Population\\ Pyramid: P3}
\subsection{Origins}
\begin{figure}
  \begin{algorithmic}[1]
  \Procedure{Iterate-P3}{}
    \State Create random $solution$
    \State Apply hill climber
    \If{$solution \notin hashset$}
      \State Add $solution$ to $P_0$
      \State Add $solution$ to $hashset$
    \EndIf

    \ForAll{$P_i \in pyramid$}
      \State Mix $solution$ with $P_i$
      \If{$solution$'s fitness has improved}
        \If{$solution \notin hashset$}
          \State Add $solution$ to $P_{i+1}$
          \State Add $solution$ to $hashset$
        \EndIf
      \EndIf
    \EndFor
  \EndProcedure
\end{algorithmic}
  \caption{One iteration of P3 optimization. $pyramid$ is an
           ordered set of populations and $hashset$ is a set
           of all solutions in $pyramid$.}
  \label{fig-p3}
\end{figure}

@@

For each problem class tested, we generated 50 instances.
Each method was run once on each instance, and limited to 3 hours of computation and 4 GB of memory.
Runs were performed on 2.5GHz Intel Xeon E5-2670v2 processors using the C++11 code available
from our website.\footnote{\url{https://github.com/brianwgoldman/GrayBoxOptimization}}
Each time the run achieved a new best fitness we recorded the current amount of processing time used.
Timing includes the discovery of subgraphs to allow for comparison between different radius values.
When reporting the ``best'' fitness for an instance we mean the best fitness found by any method before
the time limit is reached. On all Nearest Neighbor NKq instances the ``best'' fitness
is also the global optimum, verified using dynamic programming. For Ising Spin Glass the ``best'' fitness
was the global optimum 44 out of 50 times.
